---
title: "Assignment 2 - PHS 7045"
author: "Kline DuBose"
format: gfm
editor_options: 
  chunk_output_type: console
---

# Due Date
Tuesday, February 28

# Background
For this assignment, you'll be quested with speeding up some code using what you have learned about vectorization and Rcpp.

# Part 1: Vectorizing code

## Function 1
This function generates a `n x k` dataset with all its entries distributed Poisson with mean `lambda`.
```{r}
# function that was given
fun1 <- function(n = 100, k = 4, lambda = 4){
  x <- NULL
  
  for (i in 1:n) {
    x <- rbind(x, rpois(k, lambda))
  }
  
  return(x)
}

# my function
fun1alt <- function(n = 100, k = 4, lambda = 4){
  x <- matrix(data = rpois(n * k, lambda = lambda),
              ncol = k)
  return(x)
  
}

# benchmarking
bench::mark(
  fun1(),
  fun1alt(), relative = TRUE, check = FALSE
)
```

## Function 2
Like before, speed up the following functions (it is OK to use StackOverflow)

```{r}

# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n)
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  rowSums(mat)
}


# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}


fun2alt <- function(mat) {
  n <- nrow(mat)
  ans <- mat
  for (i in 1:n) {
    ans[i,] <- cumsum(mat[i,])
  }
  ans
}

## Another function I'm trying to make faster
# fun2alt <- function(mat) {
#   ans <- mat
#   ans[] <- vapply(ans, cumsum, FUN.VALUE = 1)
#   ans
# }

# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
bench::mark(
  fun1(dat),
  fun1alt(dat), relative = TRUE
)

# Test for the second
bench::mark(
  fun2(dat),
  fun2alt(dat), relative = TRUE
)
```

## Function 3
Find the column max (hint: Check out the function `max.col()`)

```{r}
#| eval: true
# Data Generating Process (10 x 10,000 matrix)
set.seed(1234)
x <- matrix(rnorm(1e4), nrow = 10)

# Find each column's max value
fun2 <- function(x) {
  apply(x, 2, max)
}

# My function
fun2alt <- function(x) {
  x[cbind(max.col(t(x)), 1:ncol(x))]
}

# Benchmarking
bench::mark(
  fun2(x),
  fun2alt(x), relative = TRUE
)

```

# Part 2: Rcpp code

As we saw in the Rcpp week, vectorization may not be the best solution. For this
part, you must write a function using Rcpp that implements the propensity score
matching algorithm. You can use [Week 5's lab](https://github.com/UofUEpiBio/PHS7045-advanced-programming/issues/8#issuecomment-1424974938) as a starting point for the problem. Your C++ file
should look something like the following:

```{Rcpp eval = FALSE}

#include<Rcpp.h>
using namespace Rcpp;

//[[Rcpp::export]]
List psmatch(
  NumericVector pscores,
  LogicalVector is_treated
)
{
  /*... setup the problem creating the output...*/
  int n = static_cast<int>(pscores.size());
  
  IntegerVector indices(n);
  NumericVector values(n);
  values.fill(std::numeric_limits< double >::max());
  
 
  /*
  ... Implement your matching (start from Week 5's lab)... 
  ... You have to consider that matches are done against groups, i.e.,
      Treated (is_treated == true) must be matched to control 
      (is_treated == false)  
  */
  
  for (int i = 0; i < n; ++1) {
    
    double & cur_best = values[i];
    auto & cur_i      = indices[i];
    
    
    
    
    
}
  
  
  // Returning
  return List::create(
    _["match_id"] = /*...*/
    _["match_pscore"] = /*...*/,
  );
}
```